---
title: "Image J Template"
output: html_document
author: Michaela Miller
email: michaela.miller@my.jcu.edu.au
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scope of the Document

This is an template to understand how to use the Image J script. The purpose of this document and code is automatically populate the size measurements based on the Raw Image J output.

To begin, you will need the following documents:
1. Microscope Sheet from the Naming Database as "Microscope Sheet.csv"
2, Raw Image J output as "Raw Image J.csv"

Please note that both sheets are required to have a column entitled "Putative.Microplastic.Sample.ID"

## Load the Packages

First, make sure the following packages are installed and loaded into the library space:
- dplyr
- readxl
- writexl
```{r load packages}
install.packages("dplyr")
library(dplyr)
install.packages("readxl")
library(readxl)
install.packages("writexl")
library(writexl)
```

## Load Your Files

Next, you'll need to upload your documents into the workspace:
```{r upload datasheets}
Raw_Image_J <- read.csv("Raw Image J.csv")
Microscope_Sheet <- read.csv("Microscope Sheet.csv")
```

## Select Appropriate Columns

In order to work with the correct data, you'll need to select certain columns from the two datasheets you've just uploaded and create new datasheets to work with, Then you'll combine these new datasheets. This will allow the original datsheets to be untouched incase a mistake occurs.
After combining, output will show you the first 10 rows to confirm this has been done properly.
```{r select columns}
Selected_Sizes <- dplyr::select(Raw_Image_J, Putative.Microplastic.Sample.ID, 
                                     Area1, Feret1, MinFeret1, Length1, Length2, Length3, Length4)
Selected_Microscope <- dplyr::select(Microscope_Sheet, Putative.Microplastic.Sample.ID, Shape)
Combined_Data <- dplyr::right_join(Selected_Microscope,Selected_Sizes)
head(Combined_Data)
```

## Do Calculations and Merge Data

Because fibres, fragments and beads are measured different as per the SOP, these three types of data will need to be treated differently. Once the calculations are finished, the data will be combined into a single new datasheet, and all unnecessary columns will be removed.
To do this we will first filter the datasheet by shape, then preform calculations and add columns for those results.
```{r calculations}
Fragments <- Combined_Data %>% dplyr::filter(Shape=="Fragment") %>% dplyr::mutate(Area=Area1+0) %>%
  dplyr::mutate(Length=Feret1+0) %>% dplyr::mutate(Width=MinFeret1+0) %>% dplyr::mutate(SD=NA)

Fibres <- Combined_Data %>% dplyr::filter(Shape=="Fibre") %>% dplyr::mutate(Area=NA) %>%
  dplyr::mutate(Length=Length1+0) %>% rowwise() %>% dplyr::mutate(Width=(Length2+Length3+Length4)/3, SD= sd(c(Length2, Length3, Length4)))

Beads <- Combined_Data %>% dplyr::filter(Shape=="Bead") %>% dplyr::mutate(Area=Area1+0) %>%
  dplyr::mutate(Length=Feret1+0) %>% dplyr::mutate(Width=MinFeret1+0) %>% dplyr::mutate(SD=NA)

Fragments_Fibres <- full_join(Fragments,Fibres) %>% dplyr::select(Putative.Microplastic.Sample.ID, Shape, Area, Length, Width, SD)
Fragments_Fibres_Beads <- full_join(Fragments_Fibres,Beads) %>% dplyr::select(Putative.Microplastic.Sample.ID, Shape, Area, Length, Width, SD)
head(Fragments_Fibres_Beads)
```

## Combine With Original Data

Now that the calculations are finished, we need to combine this sheet with the original Microscope Sheet so we can make sure all the proper data is included to then be used by either the polymer database or the R script for FTIR assignment. 
To to this we will first combine the datasheets, then rename the Shape column, and finally re-organise the columns to fit the other databases and R scripts. Again, this will give you the top rows to confirm this organisation is correct.
```{r combine with original}
Merged_Data <- dplyr::full_join(Fragments_Fibres_Beads, Microscope_Sheet, 
                                by=c("Putative.Microplastic.Sample.ID")) %>%
  dplyr::select(Putative.Microplastic.Sample.ID, Photograph.ID, Shape.x, Colour, 
                Area.x, Length.x, Width.x, SD.x, FTIR, FTIR.Mode)
Merged_Data$Shape <- Merged_Data$Shape.x
Merged_Data$Area <- Merged_Data$Area.x
Merged_Data$Length <- Merged_Data$Length.x
Merged_Data$Width <- Merged_Data$Width.x
Merged_Data$SD <- Merged_Data$SD.x
Final_Datasheet <- Merged_Data %>% dplyr::select(Putative.Microplastic.Sample.ID, Photograph.ID, 
                                                 Shape, Colour, Area, Length, Width, SD, FTIR, FTIR.Mode)
head(Final_Datasheet)
```

## Save Datasheet

The final step is to save your datasheet as "Project_Specific_Data.csv" in your workspace which can then be used for either the polymer database excel sheet or for the R Script for FTIR Assignment.
```{r save datasheet}
write.csv(Final_Datasheet, "Project_Specific_Data.csv")
```


